% !TEX root = ../msimplicial.tex

\section{Models of configuration spaces}
%We study an important example that will help us to understand how to prove our main
%theorem.

%In this section we introduce and compare multisimplicial and  simplicial  models of euclidean configuration spaces,
%related respectively to the surjection operad by McClure-Smith and the Barratt-Eccles operad.

We are interested in the homotopy type of individual euclidean configuration spaces.
Let $F_k(\R^{d})$ denote the configuration space of $k$-tuples of pairwise disjoint vectors in $\R^{d}$.
This space is equipped with a free action of the symmetric group $\sym_k$  of permutations of $\{1,\dots,k\}$ swapping elements of a $k$-tuple.

\subsection{Surjection complexes}

We define a family of multisimplicial sets
$Sur(k)$ due essentially to McClure and Smith \cite{MS}.
These have an increasing filtration by multisimplicial subsets
$Sur_d(k)$ such that the geometric realization
$\bars{Sur_d(k)}$ is homotopy equivalent to the configuration space
$F_k(\R^d)$.

%Also the geometric realizations
%$$\mathcal{F}(k)=\bars{Sur(k)}$$ appear in the work by %McClure-Smith \cite{MS}, as explained in the appendix of %\cite{Deligne}.

\begin{definition}
	Let $Sur(k)$  be the $k$-fold multisimplicial set that has
	as $(i_1,\dots,i_k)$-multisimplexes
	the surjective maps $$f:\{1,\dots,i_1+\dots+i_k+k\} \to  \{1,\dots,k\}$$ such that
	the cardinality of $f^{-1}(l)$ is $i_l$, for $l=1,\dots,k$.
	We represent such maps by the sequence
	$$f(1) \dots f(i_1+\dots+i_k+k)$$
	The face map
	$d^l_j$ removes the $(j+1)$-th occurrence of $l$ in a sequence, and the degeneracy map
	$s^l_j$ doubles the $(j+1)$-th occurrence of $l$ in a sequence.
	So for example
	\begin{align*}
		d^2_0(12321)=1321 \\ d^2_1(12321)=1231 \\ s^1_0(121)=1121
	\end{align*}
	Degenerate multisimplices are exactly the sequences containing two equal adjacent terms.

	The front  (resp. back) $(i_1,\dots,i_k)$-face of a sequence is the subsequence containing only the first (resp. last)
	$(i_l+1)$-values of $l$, for each
	$l=1,\dots,k$.
\end{definition}

The $k$-fold multisimplicial set $Sur(k)$ is filtered
by a nested family of $k$-fold multisimplicial subsets
$$Sur_0(k) \subset Sur_1(k) \subset Sur_2(k)  \subset \dots \subset Sur(k)$$
% such that $Sur(k) = {\rm colim}_n Sur_n(k)$.

A surjection belongs to $Sur_d(k)$
if any ordered subsequence with two distinct alternating values has length at most $d+1$.
For example $x=12132 \in Sur_3(3)$ but $x \notin Sur_2(3)$ because the subsequence $1212$ has length 4.

The symmetric group $\Sigma_k$ acts on $Sur(k)$ and the action preserves the filtration.

\begin{proposition} \label{sur-real}
	There is a $\Sigma_k$-equivariant homotopy equivalence
	$\bars{Sur_d(k)} \simeq F_k(\R^d)$
\end{proposition}

\begin{proof}
	The geometric realization $\bars{Sur_d(k)}$ appears in section 13 of \cite{MS} where it is denoted $Z_0^k$.
	McClure and Smith define for each $d$ a topological operads $\mathcal{D}_d$ such that $\mathcal{D}_d(k)$ is $\Sigma_k$-homeomorphic to $\bars{Sur_d(k)} \times Tot(\Delta^*)$, where $Tot(\Delta^*)$ is  contractible and equipped with trivial $\Sigma_k$-action.
	This is explained in the appendix of \cite{cyclic} where the realization is constructed as a CW complex and denoted $\mathcal{F}_d(k)$.
	McClure and Smith proceed to constructing a zig-zag of weak equivalences of operads between $\mathcal{D}_d$ and the little $d$-cubes operad
	$\mathcal{C}_d$, that in particular gives a levelwise $\Sigma_k$-equivariant equivalence $\mathcal{D}_d(k) \simeq \mathcal{C}_d(k)$ for each $k$. It is well known that $\mathcal{C}_d(k)$ is $\Sigma_k$-equivariantly homotopy equivalent to $F_k(\R^d)$ and this concludes the proof.
\end{proof}

The (normalized) chain complexes
of $Sur(k)$  $$\chi(k):=N(Sur(k))$$
were considered by McClure and Smith
\cite{MS}, %actually other paper
who constructed an operad structure
on the collections of these complexes, the {\it surjection operad} $\chi$.

The collection of chain complexes
$$\chi_d(k):=N(Sur_d(k))$$
gives a filtration of the surjection operad by suboperads
$$\chi_0 \subset \chi_1 \subset \chi_2 \subset  \dots \subset \chi$$

\begin{proposition}\label{sur-model}
	$\chi_d(k)$ and $S^{(1)}{F_k(\R^d)}$ are $\Sigma_k$-equivariantly quasi-isomorphic $E_\infty$-coalgebras
\end{proposition}

\begin{proof}
	The $\Sigma_k$-equivariant homotopy equivalence of proposition \ref{sur-real} induces a $\Sigma_k$-equivariant  quasi-isomorphism of $E_{\infty}$-coalgebras
	$$S^{(k)}(\bars{Sur_d(k)}) \simeq S^{(k)}(F_k(\R^d))$$.
	The weak equivalence %say that it is Quillen adjunction
	$Sur_d(k) \to Sing^{(k)}(\bars{Sur_d(k)})$ induces on the chain level a $\Sigma_k$-equivariant  quasi-isomorphism of
	$E_\infty$-coalgebras
	between $\chi_d(k)$
	and $S^{(k)}(\bars{Sur_d(k)})$.
	Finally the map of lemma \ref{} gives a $\Sigma_k$-equivariant quasi-isomorphism $S^{(1)}(F_k(\R^d)) \to S^{(k)}(F_k(\R^d))$ of $E_\infty$-coalgebras.
\end{proof}

\subsection{Barratt-Eccles complexes}

We recall that the Barratt-Eccles simplicial sets $\mathcal{W}\Sigma_k$ is defined levelwise by $\mathcal({W}\Sigma_k)_i=(\Sigma_k)^{i+1}$.
The face $d_{j}$ removes the $(j+1)$-st permutation, and the degeneracy $s_j$ doubles the $(j+1)$-st permutation.
There is an operad structure on the collection $\mathcal{W} \Sigma_k$.

There is a filtration of the $W \Sigma_k$ by simplicial subsets
$W_d \Sigma_k$.
Let us write elements of $\Sigma_k$ as a string containing all numbers from $1$ to $k$.
Then $(w_0,\dots,w_i) \in W_d \Sigma_k$ if for any pair of indices their order of appearance in a string swaps at most $d-1$ times, as we proceed from $w_0$ to $w_i$.
for example $y=(123,213,312) \in \mathcal{W}_3 \Sigma_3$ but $y \notin \mathcal{W}_2 \Sigma_3$
because the position of $1$ and $2$ swaps two times.

The symmetric group $\Sigma_k$ acts on
$\mathcal{W} \Sigma_k$ by diagonal action and the action preserves the filtration.

The following is well known, see for example \cite{BF}
\begin{proposition} \ref{be-real}
	The geometric realization $\bars{\mathcal{W}_d \Sigma_k}$ is $\Sigma_k$-equivariantly homotopy equivalent to $F_d(\R^k)$.
\end{proposition}

The (normalized) chain complex of $\mathcal{W}_d\Sigma_k$ is called the Barratt-Eccles chain complex $$BE_d(k):=N(\mathcal{W}_d\Sigma_k)$$
the collection of these complexes forms the Barratt-Eccles operad $BE_d$.
\cite{BF}
There is a filtration
$$BE_0 \subset BE_1 \subset BE_2 \dots \subset BE$$

Similarly as for proposition \ref{sur-model} we have the following.

\begin{proposition}
	There is a  $\Sigma_k$-equivariant quasi-isomorphism
	of $E_\infty$-coalgebras between $BE_d(k)$ and
	$S^{(1)}(F_k(\R^d))$.
\end{proposition}

\begin{proof} \label{be-model}
	The required quasi-isomorphism follows from proposition \ref{be-real} and the equivariant quasi-isomorphism $BE_d(k) \simeq S^{(1)}(|W_d\Sigma_k|)$. %counit of quillen adjunction
\end{proof}

\subsection{A comparison morphism}

In this section we construct an explicit combinatorial map between the surjection complex and the Barratt-Eccles complex, that respects filtrations, inducing weak equivalences of subcomplexes.
Furthermore it upgrades a map previously defined on the chain level by Berger and Fresse: they constructed in \cite{BFsmall}
an operad map $TR:BE \to \chi$ preserving the filtrations and a collection of $\Sigma_k$-equivariant
chain maps $TC(k):\chi(k) \to BE(k)$ preserving filtrations
that are right inverses of $TR(k)$ but do not form an operad map.
We warn the reader that neither $TR$ nor $TC$ preserve the $E_\infty$-coalgebra structure, but
$TC$ preserves the $E_1$-coalgebra structure.

We claim that $TC$ is induced by a map of simplicial sets
$$tc: Sur(k)^D \to \mathcal{W}\Sigma_k$$ that we define.
\begin{definition}
	Let $s$ be an $i$-simplex  of the diagonal $Sur(k)^D$, i.e. a sequence containing any value in $\{1,\dots,k\}$ exactly $i+1$ times.
	Then $tc(s):=(\sigma_0,\dots,\sigma_i)$ is a sequence of permutations where $\sigma_j$ is the subsequence of $s$ containing the $(j+1)$-st occurrence of each value in $\{1,\dots,k\}$.
\end{definition}

For example
$$tc(122333112)=(123,231,312)$$

\begin{proposition}
	The homomorphism $TC:\chi(k) \to BE(k)$ by Berger-Fresse satisfies
	$$TC=N(tc) \circ \ez$$ where $\ez:N(Sur(k)) \to
	N(Sur(k)^D)$ is the Eilenberg-Zilber map.
\end{proposition}

\begin{proof}
	By inspection of the definition in \cite{BFsmall}.%more? signs? which convention?
\end{proof}

%Also $\mathcal{W}$ is filtered \cite{BFsmall}, i.e.
%there is a family of nested simplicial sets %$$\mathcal{W}_1\Sigma_k \subset \mathcal{W}_2 %\Sigma_k
% \subset \dots $$ such that
%$\mathcal{W}\Sigma_k={\rm colim}_d %\mathcal{W}_d\Sigma_k$, that form a filtration by %operads
%$$\mathcal{W}_1 \subset \mathcal{W}_2 \subset \dots %\mathcal{W}$$

%The normalized chain functor defines
%$$BE_d(k):=N(\mathcal{W}_d \Sigma_k)$$ so that the %operads $BE_d$
%form a filtration of $BE$.

\subsection{Filtrations by complete graphs}
In this section we refine the filtrations of McClure-Smith and Barratt-Eccles in order to show that $tc$ induces equivariant weak equivalence on the filtation terms.

\begin{definition}[Complete graphs]
	A \textit{complete graph} on $k$ vertices is
	a pair $(\mu,\sigma)$, where $\mu$ is a collection of non-negative integers $\mu_{ij}\in\mathbb{N}$, $1 \leq i < j \leq k$
	and $\sigma$ is an ordering of
	$\{1,\dots,k\}$.
	We can draw a graph with $k$ vertices and with an oriented edge between any two vertices $i,j$
	labelled by $\mu_{ij}$, such that the orientation of the edges respect the ordering.
	For example the figure represents a complete graph on 4 vertices with ordering $\sigma=(1432)$
	\begin{equation*}
		\begin{tikzcd}
			\ & 2 & \  \\
			1 \arrow[ur,"2"] \arrow[rr] \arrow[dr,"3"']& \arrow[r,"1"] \arrow[u,"2"] & 3 \arrow[ul,"1"']  \\
			\ & 4 \arrow[ur,"4"'] \arrow[uu] & \
		\end{tikzcd}
	\end{equation*}
	Let us denote the set of complete graphs with $k$ vertices as $\mathcal{CG}(k)$.

	We write $\sigma_{ij}$ for the restriction of the ordering $\sigma$ to the set $\{i,j\}$.
\end{definition}

\begin{definition}[Complete graphs form a poset]
	$\mathcal{CG}(k)$ has a poset structure as follow:
	\begin{equation*}
		(\mu,\sigma)\le (\nu,\tau) \ \ \text{ if and only if } \ \ (\mu_{ij}<\nu_{ij}) \ \ \text{ or } \ \ (\mu_{ij},\sigma_{ij})= (\nu_{ij},\tau_{ij})
	\end{equation*}
	for each pair $\left\lbrace i,j\right\rbrace \subset\left\lbrace 1,\dots,k  \right\rbrace $.
\end{definition}

\begin{definition}[Filtration of the surjection multisimplicial set $Sur$]
	Fix a surjection $f\in Sur(k)_{i_{1},\dots,i_{k}}$.
	For any pair $(i,j)$ with $i< j$, let $f_{ij}$ be the subsequence of $f(1) \dots f(i_1+\dots+i_k+k)$ obtained omitting all the occurrences of elements different from $i$ and $j$.
	%		For example, let $31231 \in Sur(3)_{2,1,2}$, we have $(31231)_{12}=121$, $(31231)_{23}=323$ and $(31231)_{13}=3131$.
	The surjection $f$ belongs to $Sur_{(\mu,\sigma)}(k)\subseteq Sur(k)$ if for any pair $(i,j)$ with $i< j$, either in the sequence $f_{ij}$ $i$ and $j$ alternate strictly less than $\mu_{ij}$ times,
	%		, that is the amount of times that the order of $i$ and $j$ change ,
	or $i$ and $j$ alternate in the sequence $f_{ij}$  exactly $\mu_{ij}$ times and the ordering formed by the first occurrences of $i$ and $j$ in $f_{ij}$ agrees with $\sigma_{ij}$.
	Notice that \begin{equation*}
		\label{def}
		Sur_{n}(k)=\bigcup_{\max_{i<j}(\mu_{ij})< n} Sur_{(\mu,\sigma)}(k)
	\end{equation*}
	%		\\
	%		\\
	%		Returning to the example above we have $\mu_{12}(31231)=1$, $\mu_{23}(31231)=1$ and $\mu_{13}(31231)=2$ so that $31231\in Sur_{(\mu,\sigma)}(3)_{2,1,2}$
	%		\begin{itemize}
		%		\item if $\mu_{12}>1$, $\mu_{23}>1$ and $\mu_{13}>2$ and any $\sigma\in \Sigma_{3}$;
		%		\item if some $\mu_{ij}=\mu(f)_{ij}$ and the first occurences of $i$ and $j$ in $f_{ij}$ agrees with $\sigma_{ij}$.
		%		\end{itemize}
	%	    In our example we obtain that, by equation \ref{def},
	%	    $31231\in Sur_{3}(k)_{2,1,2}$
	%	    \\
	%	    \\
	% We have obtained the filtration $$Sur_1(k) \subset Sur_2(k)  \subset Sur_3(k)  \subset \dots $$
	% that induces the filtration of the correspondent diagonal simplici%al set
	%	    $$Sur_1(k)^{D} \subset Sur_2(k)^{D}  \subset Sur_3(k)^{D}  \subset\dots $$

\end{definition}

\begin{definition}[Filtration of the Barratt-Eccles simplicial set $\mathcal{W}\Sigma_k$]
	Fix an element  $w\in (\mathcal{W}\Sigma_k)_{d}$, $w=(w_{1},\dots , w_{d})$ with $w_{h}\in \Sigma_k$.
	For any pair $(i,j)$ with $i< j$, denote $w_{ij}=(w_{1,ij},\dots , w_{d,ij})$ where $w_{h,ij}$ is the subsequence of $w_{h}$ obtained omitting all the occurrences of elements different from $i$ and $j$.
	%	For example, let $(123,231,312) \in (\mathcal{W}\Sigma_3)_{3}$, we have $(123,231,312)_{12}=(12,21,12)$, $(123,231,312)_{23}=(23,23,32)$ and $(123,231,312)_{13}=(13,31,31 )$.
	The element $w$ belongs to $\mathcal{W}_{(\mu,\sigma)}\Sigma_k\subseteq \mathcal{W}\Sigma_k$ if for all pairs $(i,j)$, $i< j$ either along the sequence $w_{ij}$ $i$ and $j$ swap less than $\mu_{ij}$ times,
	%	, that is, as before,  the amount of times that the order of $i$ and $j$ change ,
	or they swap exactly $\mu_{ij}$ times and the first permutation $w_{1,ij}$ is equal to $\sigma_{ij}$.
	Notice that
	\begin{equation*}
		\label{def}
		\mathcal{W}_{n}\Sigma_k=\bigcup_{\max_{i<j} (\mu_{ij})< n} \mathcal{W}_{(\mu,\sigma)}\Sigma_{k}
	\end{equation*}
	\\
	\\
	%	Returning to the example above we have $\mu_{12}(123,231,312)=2$, $\mu_{23}(123,231,312)=1$ and $\mu_{13}(123,231,312)=1$ so that $(123,231,312)\in (\mathcal{W}_{(\mu,\sigma)}\Sigma_3)_{3}$
	%	\begin{itemize}
		%		\item if $\mu_{12}>2$, $\mu_{23}>1$ and $\mu_{13}>1$ and any $\sigma\in \Sigma_{3}$;
		%		\item if some $\mu_{ij}=\mu_{ij}(w)$ and the first permutation of $w$ is equal to $\sigma_{ij}$.
		%	\end{itemize}
	%	In our example we obtain that, by equation \ref{def},
	%	$(123,231,312)\in (\mathcal{W}_{3}\Sigma_3)_{3}$
	%	\\
	%	\\
	%	We have obtained the filtration $$\mathcal{W}_1\Sigma_k \subset \mathcal{W}_2 \Sigma_k
	%	\subset \mathcal{W}_3 \Sigma_k
	%	\subset \dots $$
	%	that is the other one we need in $tc$ definition.
\end{definition}

% In both cases, to see even the role of permutations we'll use always pairs $(\mu,\sigma)$ with $\mu$ exactly the collection of variation numbers obtained from the initial element so that the permutation is uniquely determined, and $(\mu,\sigma)$ is minimal in the poset.
% \\
% In our examples
% \begin{itemize}
	% 	\item $31231\in Sur_{(\mu,\sigma)}(3)_{2,1,2} \subseteq Sur_{3}(3)_{2,1,2}$ with $(\mu_{12},\mu_{13},\mu_{23})=(1,2,1)$ and $\sigma=(312)$
	% 	\item $(123,231,312)\in (\mathcal{W}_{(\mu,\sigma)}\Sigma_3)_{3} \subset (\mathcal{W}_{3}\Sigma_3)_{3}$ with $(\mu_{12},\mu_{13},\mu_{23})=(2,1,1)$ and $\sigma=(123)$
	% \end{itemize}
%	We observe that the simplicial map $tc$ respects the filtration, sending
%	$Sur_d(k)^D$ to $\mathcal{W}_d\Sigma_k$
%	\\
%	\begin{example}
	%	We are givine some example of the compatibility of $tc$ to visualize why it works in general.
	%	Let $122333112 \in Sur(3)^{D} $
	%	\begin{itemize}
		%		\item $(122333112)_{12}=122112\rightarrow \mu_{12}=2$
		%		\item $(122333112)_{13}=133311\rightarrow \mu_{13}=1$
		%		\item $(122333112)_{23}=223332\rightarrow \mu_{23}=1$
		%	\end{itemize}
	%So $\mu=(\mu_{12},\mu_{13},\mu_{23})=(2,1,1)$
	%and the first occurrences of $1,2,3$ give us the permutation $\sigma=(123)$ $\Rightarrow$ $122333112 \in Sur_{((2,1,1),(123))}(3)^{D}$ $\Rightarrow$ $122333112 \in Sur_{3}(3)^{D}$.
	%Now applying $tc$ we obtain $tc(122333112)=(123,231,312)\in (\mathcal{W}\Sigma_3)_{3} $
	%\begin{itemize}
	%\item $(123,231,312)_{12}=(12,21,12)\rightarrow \mu_{12}=2$
	%\item $(123,231,312)_{13}=(13,31,31)\rightarrow \mu_{13}=1$
	%\item $(123,231,312)_{23}=(23,23,32)\rightarrow \mu_{23}=1$
	%\end{itemize}
	%So $\mu=(\mu_{12},\mu_{13},\mu_{23})=(2,1,1)$
	%and the first permutation in $(123,231,312)$ give us the permutation $\sigma=(123)$ $\Rightarrow$ $(123,231,312) \in (\mathcal{W}_{((2,1,1),(123))}\Sigma_{3})_{3}$ $\Rightarrow$ $(123,231,312) \in (\mathcal{W}_{3}\Sigma_{3})_{3}$.
	%
	%
	%
	%	\end{example}

\begin{lemma}[Compatibility of $tc$]
	We have that the map $tc:Sur(k)^{D}\rightarrow \mathcal{W}\Sigma_{k}$ is compatible with the filtrations, in the sense that
	$$tc(Sur_{(\mu,\sigma)}(k)^{D})\subseteq \mathcal{W}_{(\mu,\sigma)}\Sigma_{k}$$
	and in particular
	$$tc(Sur_{n}(k)^{D})\subseteq \mathcal{W}_{n}\Sigma_{k}$$
\end{lemma}

\begin{proof}
	Let  $f\in Sur(k)^{D}$ be an $m$-simplex and denote
	$$tc(f)=w=(w_{1},\dots,w_{m+1})\in (\mathcal{W}\Sigma_{k})_{m+1}$$
	Suppose that in $f_{ij}$ $i$ and $j$ alternate $\mu_{ij}$ times, and in $w_{ij}$ $i$ and $j$ alternate  $\mu_{ij}'$ times.
	We will prove that
	$\mu_{ij}' \leq \mu_{ij}$.

	Suppose without loss of generality that $i$ occurs before $j$ in $f_{ij}$, that starts with $i$ repeated $h$ times, followed by $j$ repeated $l$ times, and
	then by $i$ again (or terminating).
	If the sequence terminates then $h=l=k$ and $\mu_{ij}=\mu_{ij}'=0$.
	For $h <l$
	let $\bar{f}_{ij}$ be the subsequence of $f_{ij}$
	obtained taking out the first $h$ values of $i$ and of $j$.
	Let $\bar{\mu}_{ij}$ be the number of variations of $i$ and $j$ in $\bar{f}_{ij}$
	and $\bar{\mu}'_{ij}$ the number of variations of $i$ and $j$ in $tc(\bar{f}_{ij})$.

	Then $\bar{f}_{ij}$
	starts with $j$,
	$\mu_{i,j}=1+\bar{\mu}_{ij}$,
	and $\mu'_{ij}=1+\bar{\mu}'_{ij}$.

	If $h \geq l$ and $l<k$
	let $\bar{f}_{ij}$ be the subsequence of $f_{ij}$ obtained taking out the first $l$ values of $i$ and $j$.
	Then $\bar{f}_{ij}$
	starts with $i$,
	$\mu_{i,j}=1+\bar{\mu}_{ij}$,
	and $\mu'_{ij}=\bar{\mu}'_{ij}$.
	By induction on the length of sequences we obtain that
	$\mu_{ij}'\le \mu_{ij}$.\\
	Moreover $tc$ is compatible with orderings: the first occurrences of $i$ and $j$ form the ordering $w_{1,ij}$ that is the first permutation of $tc(f_{ij})$.
	This concludes the proof.
\end{proof}

\begin{corollary} \label{cor-comparing}
	The map $tc_{n}:Sur_{n}(k)^{D}\rightarrow \mathcal{W}_{n}\Sigma_{k}$ is a weak equivalence.
\end{corollary}

\begin{proof}
	We can express source and target as colimits of contractible %ok?
	simplicial sets along the poset of complete graphs,
	so we have a commutative diagram
	\begin{equation*}
		\begin{tikzcd}
			hocolim_{(\mu,\sigma)}Sur_{(\mu,\sigma)}(k)^{D} \arrow[r] \arrow[d]& hocolim_{(\mu,\sigma)}\mathcal{W}_{(\mu,\sigma)}\Sigma_{k} \arrow[d] \\
			Sur_{n}(k)^{D}\arrow[r] & \mathcal{W}_{n}\Sigma_{k}&
		\end{tikzcd}
	\end{equation*}
	Where the vertical and top arrows
	are weak equivalences, and so the bottom map is a weak equivalence.
\end{proof}
We remark that $tc$ is compatible with the equivalences in propositions \ref{sur-real} and \ref{be-real} since they are proved to be equivalences by a similar method.


%Let  $F_k(\R^{d})$ be the ordered configuration %space of $k$-tuples of points in  $\R^{d}$.

%\begin{proposition}
%The geometric realizations satisfy
%$$|Sur_d(k)| \simeq |W_d(k)|  \simeq F_k(\R^{d})$$
%\end{proposition}

%\begin{proof}
%The configuration space is homotopy equivalent to %the homotopy colimit of contractible spaces along %the same poset as above, as shown by McClure and %Smith \cite{MS}. %this is originally Berger's %criterion..
%\end{proof}

\subsection{Counting  generators}

We stress that the number of generators in $\chi_d(k)$, corresponding to non-degenerate surjections,
is much smaller than the corresponding number of generators in $BE_d(k)$.
We give some examples.
Let us consider the generating polynomial functions counting the generators
$$PB_d^k(x) = \sum_i rank(BE_d(k)_i) x^i $$ $$P\chi_d^k(x)=
\sum_i rank(\chi_d(k)_i) x^i$$
Then for example
\begin{align*}
	& PB_2^4(x)=24(1+23x+104x^2+196x^3+184x^4+86x^5+16x^6)\\
	& P\chi_2^4(x)=24(1+6x+10x^2+5x^3) \\
	& \\
	& PB_3^3(x) = 6(1+5x+25x^2+60x^3+70x^4+38x^5+8x^6 ) \\
	&  P\chi_3^3(x)= 6(1+3x+7x^2+9x^3+6x^4+x^5)
\end{align*}
Therefore the multisimplicial approach using $\chi$ is much more efficient than the simplicial
approach using $BE$, when performing computations as in \cite{formality}.

