% !TEX root = ../msimp.tex

\section{Introduction}\label{s:introduction}

The cochain complex of a simplicial set is equipped with the classical Alexander--Whitney product defining the ring structure in cohomology.
This cochain level structure has several explicit extensions to an $E_\infty$-algebra \cite{mcclure2003multivariable, berger2004combinatorial, medina2020prop1} encoding commutativity and associativity up to coherent homotopies.
The importance of $E_\infty$-algebras in homotopy theory is well known.
For example, Mandell showed that finite type nilpotent spaces are weakly equivalent if and only if their singular cochains are quasi-isomorphic as $E_\infty$-algebras \cite{mandell2006homotopy_type}.
Our first objective is to define a natural product together with an $E_\infty$-algebra extension on the cochains of multisimplicial sets \cite{gugenheim1957supercomplexes}.
These are generalizations of both simplicial and cubical sets which are useful for concrete computations since they can model homotopy types using fewer cells.
For example, the proof of the non-formality of the cochain algebra of planar configuration spaces \cite{salvatore2020planarnonformality} used a simplicial model and the Alexander--Whitney product on its cochains.
By using a multisimplicial model and the product defined here, these computations become simpler and faster, paving the way for extending this result to higher dimensions.

Multisimplicial sets are contravariant functors from products of the simplex category $\simplex$ to $\Set$.
Explicitly, for any positive integer $k$ the category $\mSet{k}$ of $k$-fold multisimplicial sets is the presheaf category $\Fun((\simplex^\op)^{\times k}, \Set)$.
There is a notion of geometric realization for multisimplicial sets, which results in a CW complex having, for each non-degenerate multisimplex, a cell modeled on a product of geometric simplices $\gmsimplex{n_1}{n_k}$.
We are interested in modeling homotopy types algebraically, for which we consider the composition of the geometric realization and the functor of cellular chains $\gchains$.
This composition defines $\chains \colon \mSet{k} \to \Ch$, the functor of (normalized) chains.
In \cref{ss:e-infty extension} we define a lift of $\chains$ to the category of $E_\infty$-coalgebras, and, consequently, a lift of the functor of cochains to the category of $E_\infty$-algebras.
We do so using the finitely presented $E_\infty$-prop introduced in \cite{medina2020prop1} and its monoidal properties.
Specifically, using the isomorphism
\[
\gchains(\gmsimplex{n_1}{n_k}) \cong
\gchains(\gsimplex^{n_1}) \ot\dotsb\ot \gchains(\gsimplex^{n_k}),
\]
we extend the image of the prop generators constructed in \cite{medina2020prop1} from the chains of standard simplices to those of standard multisimplices.
These generators are the Alexander--Whitney coproduct, the augmentation map, and an algebraic version of the join product.
The resulting $E_\infty$-coalgebra structure generalizes those defined in \cite{mcclure2003multivariable, berger2004combinatorial, medina2020prop1} for simplicial chains and in \cite{medina2022cube_einfty} for cubical chains.
As an application, we study the Steenrod construction for multisimplicial chains in \cref{ss:cup coproducts} emphasizing the explicit nature of our construction.

Let us now focus on the relationship between multisimplicial and simplicial theories.
The restriction to the image of the diagonal inclusion $\simplex^\op \to (\simplex^\op)^{\times k}$ of any $k$-fold multisimplicial set $X$ defines its associated diagonal simplicial set $X^\diag$.
There is a natural homeomorphism of realizations $\bars{X} \cong \bars{X^\diag}$
\cite{quillen2010higheralgktheory}.
Under this homeomorphism the cells of $\bars{X^\diag}$ arise from those of $\bars{X}$ through subdivision, a procedure described algebraically by the Eilenberg--Zilber quasi-isomorphism
\[
\EZ \colon \chains(X) \to \chains(X^\diag).
\]
The functor induced by the diagonal restriction has a right adjoint $\cN^{(k)}$, the multisimplicial nerve of a simplicial set, making the categories of $k$-fold multisimplicial and simplicial sets equivalent in Quillen's sense.
Furthermore, there is a natural inclusion
\[
\In \colon \chains(Y) \to \chains(\cN^{(k)} Y)
\]
which is also a quasi-isomorphism.
On one hand, the $\EZ$ map preserves the counital coalgebra structure, but it does not respect the higher $E_\infty$-structure.\footnote{The Alexander--Whitney chain homotopy inverse to $\EZ$ is not a coalgebra map in general.}
On the other, the map $\In$ is an $E_\infty$-coalgebra quasi-isomorphism as proven in \cref{ss:inclusion}.
We use this fact to prove in \cref{ss:singular} that, for any topological space $\fX$, the linear map from its singular simplicial chains to its singular $k$-fold multisimplicial chains, given by precomposing a continuous map $(\gsimplex^n \to \fX)$ with the projection $(\gsimplex^n \times \gmsimplex{0}{0} \xra{\pi_1} \gsimplex^n)$, induces a natural quasi-isomorphism of $E_\infty$-coalgebras.

In the second part of the paper, we use these constructions to study a multisimplicial model of the canonical filtration
\[
\conf{r}{1} \subseteq \conf{r}{2} \subseteq \dotsb
\]
of the space $\conf{r}{\infty}$ of $r$ distinct ordered points in $\R^\infty = \colim (\R^1 \subseteq \R^2 \subseteq \dotsb)$.
Concretely, for any integer $r$, McClure and Smith \cite{mcclure2003multivariable} introduced a chain complex $\cX(r)$ of $\Z[\sym_r]$-modules with a filtration
\[
\cX_{1}(r) \subseteq \cX_{2}(r) \subseteq \dotsb
\]
and showed that $\cX(r)$ is connected to the singular chains of $\conf{r}{\infty}$ via a zig-zag of filtration preserving $\sym_r$-equivariant quasi-isomorphisms.
Presumably it was observed by both McClure--Smith and Berger--Fresse that $\cX(r)$ can be interpreted as the chains of an $r$-fold multisimplicial set $\sur(r)$, which we introduce in \cref{ss:surjection model} with a filtration
\[
\sur_{1}(r) \subseteq \sur_{2}(r) \subseteq \dotsb
\]
so that $\chains\sur_{d}(r) \cong \cX_{d}(r)$.
There is an operad structure on $\set{\cX_{d}(r)}_{r \geq 1}$ for each $d \geq 1$, but we do not focus on it since it is not induced from one at the multisimplicial level.
By the constructions in \cref{s:multisimplicial} the complex $\chains\sur(r)$ is equipped with an $E_\infty$-coalgebra structure, which we connect to the singular chains of $\conf{r}{\infty}$ via an explicit zig-zag of filtration preserving $\sym_r$-equivariant quasi-isomorphisms of $E_\infty$-coalgebras.

In a similar way, Berger and Fresse \cite{berger2004combinatorial} studied a chain complex $\cE(r)$ of $\Z[\sym_r]$-modules with a filtration
\[
\cE_{1}(r) \subseteq \cE_{2}(r) \subseteq \dotsb \,.
\]
This complex comes from the chains on a simplicial set introduced by Barratt and Eccles \cite{barrat1974operad} equipped with a filtration
\[
\BE_{1}(r) \subseteq \BE_{2}(r) \subseteq \dotsb
\]
due to Smith \cite{smith1989filtration}.
(As before we disregard the operadic structure.)
Since $\cE(r)$ is induced from a simplicial set, it is endowed with an $E_\infty$-coalgebra structure, and it is not hard to see that the zig-zag of filtration preserving $\sym_r$-equivariant quasi-isomorphisms used to compare it to the singular chains of $\conf{r}{\infty}$ respects this higher structure.
Consequently, $\cX(r)$ and $\cE(r)$ can be related by an explicit zig-zag of such maps.

It is desirable to have a direct map between the multisimplicial and simplicial models.
Berger--Fresse constructed two such filtrations preserving $\sym_r$-equivariant quasi-isomorphisms
\[
\TR \colon \chains\BE(r) \to \chains\sur(r)
\quad \text{and} \quad
\TC \colon \chains\sur(r) \to \chains \BE(r).
\]
The first one, introduced in \cite[1$\cdot$3]{berger2004combinatorial}, is unfortunately not a coalgebra map.
Therefore we will focus on the second one, which was introduced in \cite{berger2002prismatic}.
Our contribution, presented in \cref{ss:table completion}, is the construction of a factorization
\[
\TC \colon \chains\sur(r) \xra{\EZ} \chains\sur(r)^\diag \xra{\chains(\tc)} \chains \BE(r),
\]
where the second map is induced from a filtration preserving $\sym_r$-equivariant weak-equivalence of simplicial sets.
Therefore, we prove that $\TC$ is a coalgebra map since $\EZ$ is one.